### 2014-01-10 RefereeResponse/RefereeResponse_v2.tex
Added Christophs edits and committing Henrys referee response. --Lindley

\documentclass[11pt]{article}

\usepackage[usenames,dvips]{pstcol}

\usepackage{color}

\usepackage{amssymb}

\usepackage{amsmath}

%

\input{symbols.tex}

%

\newcounter{question}

\newcommand{\noteq}{{\vskip0.1in\bf \large \color{red} [Q:{\color{black}\addtocounter{question}{1}\arabic{question}}]~}}

\newcommand{\noteQ}{{\bf \LARGE \color{blue} [Q: {\color{black}\addtocounter{question}{1}\arabic{question}}]~}}

\newcommand{\notea}{{\bf \large \color{red} [A]~}}

\newcommand{\noteA}{{\bf \LARGE \color{blue} [A]~}}

\newcommand{\noteqq}[1]{{\bf \large \color{red} [Q: {\color{black}#1}]~}}

\newcommand{\noteaa}[1]{{\bf \large \color{red} [A: {\color{black}#1}]~}}

%

\textwidth 6.5in

\textheight 9.4in

\topmargin -0.8in

\oddsidemargin 0.0in

\evensidemargin 0.0in

\parindent 0.0in

\def\baselinestretch{1.0}

\def \Russ {Translated from Russian. Fixed. Thanks.}

\def \F    {Fixed.}

\def \FT   {Fixed. Thanks.}

\def \D    {Done.}

\def \AP   {Basically Stylistic- Author's Prerogative.}

\begin{document}

\begin{flushright}

\today

\end{flushright}

%----------------------------------------------------------------

Dear JINST Editors,\\

\parindent 1.0in

We thank the referees for their very helpful and knowledgeable

comments. We have substantially rewritten the manuscript in response

to the comments as detailed in the list below.

There are several areas, however, in which the referees requested

substantially expanding the scope of the paper.  We envisioned this

paper as a proof of  the principle that the use of the fast timing

photodetectors presently being developed would enable the separation

of Cherenkov and scintillation light, as the reconstruction of the

direction of low energy ($\sim$1 MeV) electrons in a liquid

scintillator detector has never been shown before.  In this paper, we

want to: 1) show scintillation/Cherenkov separation is possible; 2)

determine the parameters of a large scintillator detector that are

important to this separation; and 3) demonstrate that reconstruction

techniques based on Cherenkov light would work in idealized

situations.

We are now working on two follow-up papers that are substantially

beyond the scope of the present paper, moving beyond

proof-of-principle into the laborious and time-intensive area of

reconstruction algorithms, areas for which the difficulty is known

well by some of us from our experience on KamLand. The first is a

detailed study of reconstruction in more realistic scenarios. The

second is a proposal for a $0\nu\beta\beta$ experiment  where the detailed

background reduction factors and sensitivity to new physics are

presented, SuperNEMO has a similar paper, but for scintillator

detectors nothing similar exists since a directional signal has not

been considered for them in the past.  Finally, we will be having a

turnover of young personnel so this paper provides an accessible summary

of their work.

\parindent 3.0in

    Sincerely,\\

\parindent 4.0in

		         Lindley Winslow, for the Authors

\parindent 0.0in

%\begin{verbatim}

%\end{verbatim}

\noteq [Referee 1 Intro] ``I like  very much the clean, transparent, simple

and well-motivated methodology (with one exception..). The structure

of the paper is nice''

\notea{\bf Many thanks.}\\

\noteq [Referee 1 Intro] ``However the presentation of the results is a

different thing... I am particularly concerned with the fact...,

instead of showing the resulting CosTheta distribution that would seem

like the obvious result one wants to see. .... Why not to disclose

this openly to allow the reader to evaluate the results in detail?''

\notea{\bf  This is our fault- the desired plot was (is) there

(Figure 3), but was labeled as $p_x/|\vec{p}|$ rather than the obvious

CosTheta that it is. We have made this clear on the figure and in the

text. Thanks.}\\

\noteq [Referee 1: Page 1] ``Please specify FWHM or sigma and give a

reference for both cases.''

\notea{\bf  \D }\\

DOES ANYBODY KNOW WHAT HSE IS ASKING FOR WRT THE REFERENCE?

\noteq [Referee 1: Page 1] ``...but please avoid misplacing the adjective good in

this context. ...''

\notea{\bf  \FT}\\

\noteq [Referee 1: Page 2] ``..the authors should..check and refer to

the work of NEXT and Super-Nemo collaborations:''

\notea{\bf  \D}\\

I DON'T SEE THE REFERENCE TO NEXT IN THE TOPOLOGY PARAGRAPH- DID I MISS

IT? IT SHOULD GO AFTER ``..ESPECIALLY FOR 0NUBETABETA SEARCHES'''?

\noteq [Referee 1: Page 2] ``Define PPO here, not later''

\notea{\bf \D}\\

\noteq [Referee 1: Page 3] ``Please define all magnitudes in eq. 2.1 and 2.2''

\notea{\bf \D}\\

\noteq [Referee 1: Page 3]  ``Indicating the direction of the

electrons should not harm''

\notea{\bf  \D}\\

WAS ANYTHING CHANGED IN THE FIGURE? THE CAPTION MAKES IT CLEAR, AND

THE BRIGHT WHITE AXIS ARROW IS CLEAR... IF WE DON'T CHANGE THE FIG, WE

SHOULD SAY SO AND WHY NOT (EASY)

\noteq [Referee 1: Page 3] ``None of the best limits or claims in bb0

sues an isotope that satisfies this condition..''

\notea{\bf \FT}\\

IS THIS IS FIXED? I COULDN'T FIND IT IN THE CURRENT DRAFT AROUND

THE 2.6 MEV DISCUSSION. .I'M CONFUSED BY THE COMMENT, FRANKLY. DOES

THE PRESENT TEXT NEED MORE WORK? (I.E. I NEED EDUCATING...)

\noteq [Referee 1: Page 3]

\notea{\bf \FT}\\

THIS IS FIXED, RIGHT? I COULDN'T FIND IT IN THE CURRENT DRAFT AROUND

THE 2.6 MEV DISCUSSION. .I'M CONFUSED BY THE COMMENT, FRANKLY. DOES

THE TEXT NEED WORK? (I.E. I NEED EDUCATING...)

\noteq [Referee 1: Page 4] ``could you perhaps include the

attenuation @400nm and n@400..?''

\notea{\bf \D}\\

I SEE THE ATTENUATION LENGTH, BUT NOT N@400

\noteq [Referee 1: Page 4] Mentioning the figure in the citation.

\notea{\bf ?}\\

I DIDN'T UNDERSTAND THIS COMMENT- THE PAPER MENTIONS FIG 4C..

\noteq [Referee 1: Page 4] ``Could you please include also the cutoff

wavelength previously referred to?''

\notea{\bf ?}\\

I DIDN'T UNDERSTAND THIS ONE EITHER

\noteq [Referee 1: Page 4] `` Scattering have not been included..''

\notea{\bf We have added a figure and accompanying text to show that

  the scattering only adds a late tail. Thanks.}\\

\noteq [Referee 1: Page 4] ``For which digitized values come from ??''

\notea{\bf \F}\\

\noteq [Referee 1: Page 4] ``5 MeV electrons.. It is very disturbing

to benchmark the simulation with electrons having an energy doubling

the Qbb of the most common isotopes..''

\notea{\bf We have clarified in several places

that the 5 MeV case is for neutrino-electron experiments. For dbb we

have made it explicit that we consider 1.4 and 2.1 MeV}\\

\noteq [Referee 1: Page 4] ``Please specify also the track length in

the direction of emission''

\notea{\bf \D}\\

\noteq [Referee 1: Page 5] ``Before the solutes in liquid... Please rephrase''

\notea{\bf \D}\\

\noteq [Referee 1: Page 5] ``I hope that [43] is not the final version

but a draft.

\notea{\bf It was a draft; we have changed the reference. Thanks!}\\

\noteq [Referee 1: Page 6] ``So I would be very much convinced ..if

the authors would report Fig. 3 during the electron history

(equivalently, energy or distance, including several histograms for

each case. It would be nice to see at which energy scale or length the

electron isotropizes and the Cherenkov cone smears out.''

\notea{\bf We have added studies of 1.4 MeV and 2.1 MeV electrons,

  which show that the separation and directionality based on timing

  hold as the electron energy is lowered, as shown in the newly-added

  Figures 5 and 6. We have also added the path

  lengths and total distance and time from the vertex. We believe that

  additional plots beyond the ones we have added won't add any useful

  information.}\\

 HSE'S CLEARLY STUCK ON THIS- MISSED THE POINT. HOWEVER, WE NEED TO

  ANSWER THIS DIRECTLY - IS IT POSSIBLE TO MAKE ONE PLOT OF

\noteq [Referee 1: Pages 7-8] ``This is interesting, but specially

section 6 seems a bit off-topic to me. I am not against keeping it...''

\notea{\bf We'd like to keep it. }\\

\noteq [Referee 1: Page 9] ``Please add a histogram of the reconstructed

angle for all energies...''

\notea{\bf }\\

NEEDS ATTENTION...

\noteq [Referee 1: Page 9] ``..the vertex reconstruction has to be

obtained from all photons and the direction from the Cherenkov photons..''

\notea{\bf }\\

I CAN'T WRITE AN ANSWER TO THIS- CHRISTOPHE,ANDREY, MATT, LINDLEY?

(there's lots of email that contains the answer, it seems...)

\noteq [Referee 1: Fig. 7] ``the authors should show the spread of the

cos(theta) distribution for each of the energies.

\notea{\bf \D See Figure 3.}\\

I DON'T HAVE A FIGURE 7. ???

\noteq [Referee 1: Fig. 7 ``It would be nice to show/indicate..also the

Cherenkov threshold for high wave-length (n about 1.4)

\notea{\bf We have added this to the text.}\\

I DON'T HAVE A FIGURE 7. ???

\noteq [Referee 1: Conclusions] ``Should be rewritten after the

previous objections are addressed.''

\notea{\bf \D Thanks!}\\

WE SHOULD REWRITE THE CONCLUSIONS TO BE MORE SPECIFIC WITH MORE

FACTS. AS THEY STAND THEY'RE SORT OF GENERIC- IF THAT'S ALL ONE READS

OF THE PAPER ONE'S NOT MUCH WISER ...

\noteq [Referee 1: Page 7-8] ``I see no convincing proof of it with

the information here presented. Hope the answers help clarifying the issue.''

\notea{\bf We have made the sentence more specific. Thanks.}\\

WE SHOULD BE MORE SPECIFIC- CAN GIVE NUMBERS RATHER THAN OPINIONS

HERE..

%\newpage

\noteq [Referee 2: General Comments] ``The paper introduces an

original technical development which could increase the interest of

large scale scintillation detectors for neutrino physics in general

and neutrinoless double beta decay in particular''

\notea{\bf Warms the cockles of our hearts...}\\

\noteq [Referee 2: General comments] `` ..however the discussion

should deserve...a deeper analysis.''

\notea{\bf Yes. We have extended the analysis to lower energies and

  have extracted much more detail from the simulation, and have

  extensively rewritten the manuscript. It is still at the

  `proof-of-principle level, as to go deeper yet will require vertex

  reconstruction, which is a much bigger effort (we are working on it).}\\

\noteq [Referee 2: General Comments] ``The physics case is included in

the title and mentioned in the text but never discussed in some

detail...A description of why the method is powerful for double beta

decay, and..to what extent it can improve the experimental sensitivity

would complete the discussion

\notea{\bf }\\

NYGREN'S RIGHT- CAN WE ADD A SENTENCE OR TWO TO THE INTRO AND

(PERHAPS) THE CONCLUSIONS?

\noteq [Referee 2: General Comments] ``The chosen example (5 MeV,

head-to-head single electrons) is of little interest for double beta decay.''

\notea{\bf We have added plots and numbers for 1.4 and 2.1 MeV, and

  substantially rewritten the text to differentiate between the cases

  of neutrino-electron scattering and double-beta decay.}\\

\noteq [Referee 2: General Comments] ``The dependence on the decay

position.. is not discussed.''

\notea{\bf To do a credible and useful job on the position dependence

  will require a fairly-detailed vertex reconstruction. This is beyond

  the scope of this present `proof-of-principle' paper. We have tried

  to clarify this in the text. }\\

\noteq [Referee 2: General Comments] Threshold effects as well as the

effect of statistical fluctuations for the low energy tails.. are not

discussed. The same holds for position and direction reconstruction

capabilities''

\notea{\bf }\\

I DON'T KNOW WHAT TO SAY HERE- SOMEBODY MORE EXPERT?

\noteq [Referee 2: Abstract] ``While appropriate for neutrino physics,

energy resolution is definitely worse with respect to other techniques..''

\notea{\bf We have fixed the text. Thanks.}\\

\noteq [Referee 2: Introduction] ``A reference ... would help to understand

the connection with Majorana neutrinos.''

\notea{\bf \D}\\

\noteq [Referee 2: Page 1] Is this 1 sigma or FWHM?

\notea{\bf It's sigma- we have fixed the text.}\\

\noteq [Referee 2: Page 2] ``better than water Cherenkov detectors''

\notea{\bf \FT}\\

PLEASE ADD `` for neutrino detectors'' after ``the other developed

technology'' AT THE TOP OF PAGE 2.

\noteq [Referee 2: Page 2 ] ``strong suppressant of backgrounds''..this

statement should deserve a deeper quantitative discussion ..''

\notea{\bf We agree. However, a useful ``quantitative discussion''

  will require a vertex reconstruction. We will get there, but at present

  it's beyond the scope of `proof-of-principle, looking to see if we

  can see the signal. We have tried to make this clearer in the text. }\\

\noteq [Referee 2: Equ. 2.1] ``..the meaning of the parameter `x' is

not described.''

\notea{\bf \F}\\

\noteq [Referee 2: Page 3] What is the typical cutoff value and the

corresponding useful fraction of Cherenkov light?

\notea{\bf We have added the cutoff values.}\\

DO WE HAVE THE USEFUL FRACTIONS OF CHERENKOV LIGHT? (I DON'T EVEN KNOW

IF THAT'S WELL-DEFINED?)

\noteq [Referee 2: Page 3] `` `Probable case being...' Events with

asymmetric electron energies are far more abundant.''

\notea{\bf \FT}\\

WE SHOULD REWRITE THIS SENTENCE ..E.G. ``We consider the case of

equal..'' OR SOMETHING.. (?) (MAYBE CHANGE 'PROBABLE' TO 'SIMPLE'?)

\noteq [Referee 2: Page 4] `` `Fully absorbing' ''. How are results

influenced by deviations from this hypothesis?''

\notea{\bf This is a level of detail well beyond what we've done so

  far; once we have a real detector design and a full simulation it

  can be addressed, but at this point devoting effort to it

doesn't seem productive.}\\

\noteq [Referee 2: Page 4] ``Single 5 MeV electrons''.

\notea{\bf \F}\\

\noteq [Referee 2: Page 4] ``I believe the `realistic' analysis

mentioned here is indispensable ..for this process''

\notea{\bf We agree, and are working on it. It isn't easy, or,

  unfortunately, fast compared to a post-doc tenure.}\\

\noteq [Referee 2: Page 6] ``The dependence on the position and the

corresponding transit time variations cannot in my opinion be neglected.''

\notea{\bf }\\

THIS NEEDS AN EXPERT'S ANSWER...

\noteq [Referee 2: Figure 3] `` This figure is very interesting

to understand the capacity of the method to maintain directionality

information... The comparison with a single event (or equivalently the

distribution of the variances) and the distribution for 1 MeV

electrons would improve to understand also the limits/power of the method''

\notea{\bf }\\

THIS NEEDS AN EXPERT'S ANSWER...

\noteq [Referee 2: reconstruction] ``It is not clear if limiting the

reconstruction to Cherenkov photons improve reconstruction''

\notea{\bf Yes, we agree. We are not proposing throwing away the

scintillation light, but in addition using the selected Cherenkov light to

determine the electron direction, i.e., not `limiting the

reconstruction to Cherenkov photons''}\\

CAN THIS ANSWER BE IMPROVED? (YES)

\end{document}

\noteq [Referee 1]

\notea{\bf }\\